---
title: "ATL Hawks Analytics Final Code"
author: "Michael Lewis-Wedderburn"
date: "2023-06-04"
output: pdf_document
---

```{r}

library("dplyr") #Using dplyr library to use other functions

```


This reads the entire CSV file and attaches it's features.

```{r}

NBA <- read.csv("nbaplayersdraft.csv")
View(NBA) 
attach(NBA) #Allows the use of the features in the csv file
```

This filters the entire data set to become all players who went to Duke
and were drafted in or before the 2000 draft.

```{r}

NBA2 <- subset(NBA, college == "Duke" & year <= 2000) #Filters a new dataset
#View(NBA2)
NBA2 %>% count(team)

```

This filters the entire data set to become all players who have a first
name that begins with D and were drafted in an even year draft.

```{r}

NBA3 <- subset(NBA, year %% 2 == 0) 
NBA3 <- NBA3 %>% filter(startsWith(player, "D")) #Filters a new dataset
#View(NBA3)
NBA3 %>% count(team)

```

This filters every team's first round pick selections and plots them in
one graph.

```{r}

#For loop creates 38 separate scatterplots for each team's FR picks
for (team_name in unique(NBA$team)) {
  
  FRP <- subset(NBA, team == team_name & overall_pick <= 30)
  FRP_plot <- plot(FRP$overall_pick ~ FRP$year, main = team_name)
  
}

```

This generates the expected Value Over Return for each draft pick possible.
```{r}

#For loop assigns draft value for picks 1-60 based on E(VOR) formula
for (pick in 1:60) {
  
  Draft_Value = 21.930875 - 5.414771 * log(pick)
  print(paste("Value of Pick", pick, ":", Draft_Value))

}

```

This finds the total VOR for each NBA team and puts it into a dataframe to 
compare performance. 
```{r}

#Lists used to create dataframe
total_list = list()
NBA_team_list = list()

#For loop generates total VOR for each team and assigns them to their 
#respective lists
for (team_name2 in unique(NBA$team)) {
  
  NBA_VOR <- subset(NBA, team == team_name2)
  total_vor <- sum(NBA_VOR$value_over_replacement, na.rm=TRUE)
  #print(paste("Total VOR for", team_name2, ":", total_vor))
  total_list[[length(total_list) + 1]] = total_vor
  NBA_team_list[[length(NBA_team_list) + 1]] = team_name2
  
}

#Creates dataframe to show the value of Total VOR for each team
NBA_VOR_df <- data.frame(total_list)
colnames(NBA_VOR_df) <- paste0(NBA_team_list)    
head(sort(NBA_VOR_df, decreasing=TRUE))


```

This finds the average VOR for each college team and puts it into a dataframe to 
compare performance. 
```{r}

#Lists used to create dataframe
avg_list = list()
college_team_list = list()

#For loop generates average VOR for each team and assigns them to their 
#respective lists
for (team_name3 in unique(NBA$college)) {
  
  NCAA_VOR <- subset(NBA, college == team_name3)
  avg_vor <- mean(NCAA_VOR$value_over_replacement, na.rm=TRUE)
  #print(paste("Total VOR for", team_name3, ":", avg_vor))
  avg_list[[length(avg_list) + 1]] = avg_vor
  college_team_list[[length(college_team_list) + 1]] = team_name3
}

#Creates dataframe to show the value of Average VOR for each team
NCAA_VOR_df <- data.frame(avg_list)
colnames(NCAA_VOR_df) <- paste0(college_team_list)    
head(sort(NCAA_VOR_df, decreasing=TRUE))


```

References:
<https://www.digitalocean.com/community/tutorials/r-read-csv-file-into-data-frame>
<https://dplyr.tidyverse.org/reference/count.html>
<https://stackoverflow.com/questions/43227088/how-can-i-apply-dplyrs-select-starts-with-on-rows-not-columns>
<https://stackoverflow.com/questions/19761899/unique-values-in-each-of-the-columns-of-a-data-frame>
<https://nycdatascience.com/blog/student-works/data-analysis-on-the-value-of-an-nba-draft-pick/>
<https://www.bruinsportsanalytics.com/post/nba_first_round>
<https://www.projectpro.io/recipes/append-output-from-for-loop-list-r>
